When, why, and how to avoid, prefer, or tolerate inbreeding

### 2018-03-02 presentation/inbreeding.Rmd
More progress

- Inbreeding avoidance is predicted given a sufficiently high threshold of inbreeding depression

- Inbreeding depression thresholds differ between females and males causing sexual conflict

@@

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
Parker (1979) *Sexual selection and reproductive competition in insects* (pp. 123–166). Parker (2006) *Philos. Trans. Royal Soc. B.* [**361**:235-259](https://doi.org/10.1098/rstb.2005.1785). <br> Duthie, AB & JM Reid (2015) *PLoS One* [**10**:e0125140](https://bradduthie.github.io/Duthie%26Reid2015.pdf). 
</span></div>

@@

- Inbreeding tolerance increases with lower encounter rates (Kokko & Ots 2006)

- Dispersal has effects (Lehmann & Perrin 2003)

- Thresholds change given subsequent mating decisions of relatives (Duthie & Reid 2015)

- Thresholds change given modulation of parental investment (Duthie et al. 2016)

@@

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
Parker (1979) *Sexual selection and reproductive competition in insects* (pp. 123–166). Parker (2006) *Philos. Trans. Royal Soc. B.* [**361**:235-259](https://doi.org/10.1098/rstb.2005.1785). <br> Duthie, AB & JM Reid (2015) *PLoS One* [**10**:e0125140](https://bradduthie.github.io/Duthie%26Reid2015.pdf). 
</span></div>

@@

## Inbreeding depression ($\delta$) slopes ($\beta$)

@@

## Individual-based modelling

- Much different assumptions
- Stochasticity
- See if we can get a similar answer

## Individuals

@@

## Individuals

@@

## General individual-based model assumptions

- Non-overlapping generations

@@

>>>>

- Inbreeding avoidance is predicted given a sufficiently high threshold of inbreeding depression [1,2]

- Inbreeding depression thresholds differ between females and males causing sexual conflict [1,2]

@@

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
[1] Parker (1979) *Sexual selection and reproductive competition in insects* (pp. 123–166). <br>
[2] Parker (2006) *Philos. Trans. Royal Soc. B.* [**361**:235-259](https://doi.org/10.1098/rstb.2005.1785). 
</span></div>

@@

- Inbreeding tolerance increases with lower encounter rates [1].

@@

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
[1] Kokko, H & I Ots (2006) *Evolution* [**60**:467-475](http://onlinelibrary.wiley.com/doi/10.1111/j.0014-3820.2006.tb01128.x/abstract).
</span></div>

- Inbreeding tolerance increases with lower encounter rates [1].

- Inbreeding avoidance or preference can affect dispersal [2].

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
[1] Kokko, H & I Ots (2006) *Evolution* [**60**:467-475](http://onlinelibrary.wiley.com/doi/10.1111/j.0014-3820.2006.tb01128.x/abstract).
[2] Lehmann, L & N Perrin (2003) *Am. Nat.* [**162**:638-652](https://www.journals.uchicago.edu/doi/10.1086/378823). <br>
</span></div>

- Inbreeding tolerance increases with lower encounter rates [1].

- Inbreeding avoidance or preference can affect dispersal [2].

- Thresholds change given mating decisions of avoided relatives [3].

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
[1] Kokko, H & I Ots (2006) *Evolution* [**60**:467-475](http://onlinelibrary.wiley.com/doi/10.1111/j.0014-3820.2006.tb01128.x/abstract).
[2] Lehmann, L & N Perrin (2003) *Am. Nat.* [**162**:638-652](https://www.journals.uchicago.edu/doi/10.1086/378823). <br>
[3] Duthie, AB & JM Reid (2015) *PLoS One* [**10**:e0125140](https://bradduthie.github.io/Duthie%26Reid2015.pdf). 
</span></div>

- Inbreeding tolerance increases with lower encounter rates [1].

- Inbreeding avoidance or preference can affect dispersal [2].

- Thresholds change given mating decisions of avoided relatives [3].

- Thresholds change if parental investment can be modulated [4].

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
[1] Kokko, H & I Ots (2006) *Evolution* [**60**:467-475](http://onlinelibrary.wiley.com/doi/10.1111/j.0014-3820.2006.tb01128.x/abstract).
[2] Lehmann, L & N Perrin (2003) *Am. Nat.* [**162**:638-652](https://www.journals.uchicago.edu/doi/10.1086/378823). <br>
[3] Duthie, AB & JM Reid (2015) *PLoS One* [**10**:e0125140](https://bradduthie.github.io/Duthie%26Reid2015.pdf). 
[4] Duthie, AB et. al (2016) *Proc. R. Soc. B.* [**282**:20161845](https://bradduthie.github.io/Duthie%26Reid2015.pdf).
</span></div>

- Inbreeding tolerance increases with lower encounter rates [1].

- Inbreeding avoidance or preference can affect dispersal [2].

- Thresholds change given mating decisions of avoided relatives [3].

- Thresholds change if parental investment can be modulated [4].

- Thresholds change if there are male opportunity costs [5]

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
[1] Kokko, H & I Ots (2006) *Evolution* [**60**:467-475](http://onlinelibrary.wiley.com/doi/10.1111/j.0014-3820.2006.tb01128.x/abstract).
[2] Lehmann, L & N Perrin (2003) *Am. Nat.* [**162**:638-652](https://www.journals.uchicago.edu/doi/10.1086/378823). <br>
[3] Duthie, AB & JM Reid (2015) *PLoS One* [**10**:e0125140](https://bradduthie.github.io/Duthie%26Reid2015.pdf). 
[4] Duthie, AB et. al (2016) *Proc. R. Soc. B.* [**282**:20161845](https://bradduthie.github.io/Duthie%26Reid2015.pdf). <br>
[5] Waser PM, et. al. (1986) *Am. Nat.* [**128**:529-537](https://www.journals.uchicago.edu/doi/abs/10.1086/284585).
</span></div>

@@

- Focuses on selection to the exclusion of drift, mutation, immigration

![](images/real_pedigree.png)

- Focuses on selection to the exclusion of drift, mutation, immigration

- Does not consider realistic distributions of kinship

![](images/polyandry_eg.png)

- Focuses on selection to the exclusion of drift, mutation, immigration

- Does not consider realistic distributions of kinship

- Does not consider complex strategies such as female multiple mating (polyandry) [but see 1]

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
[1] Lehtonen, J & H Kokko (2015) *Proc. R. Soc. B* [**282**:20142716](http://rspb.royalsocietypublishing.org/content/282/1801/20142716).
</span></div>

![](images/polyandry_eg.png)

- Focuses on selection to the exclusion of drift, mutation, immigration

- Does not consider realistic distributions of kinship

- Does not consider complex strategies such as female multiple mating (polyandry) [but see 1]

- Does not account for post-copulatory inbreeding avoidance or preference

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
[1] Lehtonen, J & H Kokko (2015) *Proc. R. Soc. B* [**282**:20142716](http://rspb.royalsocietypublishing.org/content/282/1801/20142716).
</span></div>

## Individual-based modelling approach

![](images/ibm_approach.png)

- Models discrete individual females and males in a finite population

- Explicitly tracks pedigree to calculate kinship between individuals

- Allows for simultaneous evolution of inbreeding behaviour and polyandry

- Allows for the evolution of pre- and post-copulatory inbreeding behaviour

<!--- 
## Inbreeding depression ($\delta$) slopes ($\beta$)

@@

--->

## The polyinbreeding team

![](images/polyinbreeders.png)

## Individual traits and genomes

@@

## Individual traits and genomes

@@

@@

## General modelling steps
